Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-04-13 ltx/BrainIDE.tex
revived section on frequency analysis after it was lost in a conflict

Considering the spectral representation of the dynamic field at time $t$ denoted by $W_t(\boldsymbol{\nu})$ \todo{can this be $V_t(\boldsymbol{\nu})$}, multi-dimensional sampling and reconstruction theory dictates that for and accurate representation, $W_t(\boldsymbol{\nu})$ need to be spatially bandlimited. Such a condition is usually satisfied due to the bandlimiting effect of most practical kernels and the spatially colored covariance functions \todo{~cite\{some eeg literature here\}}. In addition, the Gaussian pick-up of the sensors provides a low-pass filtering effect, further limiting the bandwidth of the field. Thus the spectral response $W_t(\boldsymbol{\nu})$ is usually band-limited with $W_t(\boldsymbol{\nu}) \approx 0 ~ \forall \boldsymbol{\nu} > \boldsymbol{\nu}_c$, where $\boldsymbol{\nu}_c$ is a cutoff frequency. Note that for the 3-dimensional, homogeneous and isotopic fields being considered $\boldsymbol{\nu}_c = [\nu_c ~ \nu_c ~ \nu_c]^\top$. Given such a band-limited field and based on Shannon's sampling theorem, the distance between adjacent sensors needs to within $\frac{1}{2\nu_c}$. Satisfying this condition ensures that the hidden dynamic field $w_t(\mathbf{r})$ can be inferred from the sampled observation $\mathbf{y}(t)$.

There are two options available to alleviate problems with aliasing. The first is to increase the spatial sampling frequency (using a higher resolution of sensors), and the second is to use wider sensors, which act like an anti-aliasing filters. The first option is more desirable, since more information is retained. If sensors are modelled as having a Gaussian pickup range than the slope of the attenuation will be gradual allowing for some low amplitude high frequency components to still be present and mild aliasing will still occur. In practice, the bandwidth of the cortex can not be estimated by traditional electrophysiological measurements due to spatial limitations. It may be estimated using other modalities with higher spatial resolution such as, fMRI, NIRS or other optical imaging techniques. Errors associated with mild aliasing result in distortions in the high frequency components of the field dynamics. However, after applying the basis function decomposition, high frequency component contribute negligibly to the inverse problem and are therefore not a major problem~\cite{Sanner1992}. This is further illustrated our results.

Similar considerations need to be made regarding the representation of the dynamic field $w_t(\mathbf{r})$ by its basis functions decompositions. The minimum distance between adjacent basis functions needs to again satisfy Shannon's sampling theorem. Thus the minimum distance between basis must again satisfy $\frac{1}{2\nu_c}$. For the Gaussian basis functions $\phi(\mathbf{r})$ being considered, the basis width parameter $\sigma^2$ can also be inferred from spectral considerations \cite{Sanner1992,Scerri2009}. The Fourier transform of a Gaussian is another Gaussian with width $\sigma^2_{\nu}$ given by $\sigma^2_{\nu} = \frac{1}{\pi^2\sigma^2}$. Thus to ensure that the basis functions can represent fields with frequency content up to $\boldsymbol{\nu}_c$, $\sigma_{\nu}$ should be set with the 3dB frequency for the basis functions equal to $\boldsymbol{\nu}_c$, thus obtained $\sigma^2_{\nu}$ and therefore $\sigma^2$.

>>>>

Considering the spectral representation of the dynamic field at time $t$ denoted by $W_t(\boldsymbol{\nu})$ (\todo{can this be $V_t(\boldsymbol{\nu})$}), multi-dimensional sampling and reconstruction theory dictates that for and accurate representation, $W_t(\boldsymbol{\nu})$ need to be spatially bandlimited. Such a condition is usually satisfied due to the bandlimiting effect of most practical kernels and the spatially colored covariance functions \todo{~cite\{some eeg literature here\}}. In addition, the Gaussian pick-up of the sensors provides a low-pass filtering effect, further limiting the effective bandwidth of the field. Thus the spectral response $W_t(\boldsymbol{\nu})$ is usually band-limited with $W_t(\boldsymbol{\nu}) \approx 0 ~ \forall \boldsymbol{\nu} > \boldsymbol{\nu}_c$, where $\boldsymbol{\nu}_c$ is a cutoff frequency. Note that for the 3-dimensional, homogeneous and isotopic fields being considered $\boldsymbol{\nu}_c = [\nu_c ~ \nu_c ~ \nu_c]^\top$. Given such a band-limited field and based on Shannon's sampling theorem, the distance between adjacent sensors, $\Delta_s$, needs to satisfy
\begin{equation}
\Delta_s = \frac{1}{2\rho\nu_c},
\end{equation}
where $\rho \in \mathbb{R} \ge 1$ is an oversampling parameter. Satisfying this condition is required for the hidden dynamic field $w_t(\mathbf{r})$ to be precisely inferred from the sampled observations, $\mathbf{y}(t)$. If the condition is not satisfied, spatial aliasing will occur.

There are two options available to alleviate problems with aliasing. The first is to increase the spatial sampling frequency (using a higher resolution of sensors), and the second is to use wider sensors, which act like an anti-aliasing filters. The first option is more desirable, since more information is retained. However, if sensors are modelled as having a Gaussian pickup range than the effective filter will not have a sharp cut-off and high frequency components will still remain and mild aliasing will still occur. In practice, the bandwidth of the cortex can not be estimated by traditional electrophysiological measurements due to spatial limitations. It may be estimated using other modalities with higher spatial resolution such as, fMRI, NIRS or other optical imaging techniques. Errors associated with mild aliasing result in distortions in the high frequency components of the field dynamics. However, after applying the basis function decomposition, high frequency component contribute negligibly to the inverse problem and are therefore not a major problem for the state and parameter estimation~\cite{Sanner1992}. This is further illustrated our results.
$\mathcal{W}$
Similar considerations need to be made regarding the representation of the dynamic field $w_t(\mathbf{r})$ by its basis functions decompositions. The minimum distance between adjacent basis functions also needs to satisfy Shannon's sampling theorem. Thus, the minimum distance between basis must satisfy
\begin{equation}
\Delta_b < \frac{1}{2\nu_c}.
\end{equation}
For the Gaussian basis functions $\phi(\mathbf{r})$ being considered, the basis width parameter $\sigma^2$ can also be inferred from spectral considerations \cite{Sanner1992,Scerri2009}. The Fourier transform of a Gaussian is another Gaussian with width $\sigma^2_{\nu}$ given by
\begin{equation}
\sigma^2_{\nu} = \frac{1}{\pi^2\sigma^2}.
\end{equation}
Thus to ensure that the basis functions can represent fields with frequency content up to $\boldsymbol{\nu}_c$, $\sigma_{\nu}$ should be set with the 3~dB frequency for the basis functions equal to $\boldsymbol{\nu}_c$, thus obtained $\sigma^2_{\nu}$ and therefore $\sigma^2$.
